---
title: Computational Notebook
subtitle: Main Analysis for BP test-retest reliability
date: today
format:
  html:
    toc: true
    toc-location: right-body
    toc-depth: 3
    toc-title: Contents
    embed-resources: true
    self-contained-math: true
    code-fold: true
    theme: lumen
    number-sections: false
    #number-depth: 2

---

```{r Set-Up}
#| output: false

library(dplyr)
library(purrr)
library(tidyr)
library(ggplot2)
library(rstatix)
library(SimplyAgree)
library(boot.pval)
library(tidytext)
library(scales)

set_theme(theme_bw())
```

# Background

A great deal of research interest has focused on identifying breakpoints in NIRS-derived measures. Some studies report no significant difference between methods, while others find breakpoints at different workloads. A key challenge is that even when the mean difference between methods is small, individual-level agreement is often poor. This poor individual-level agreement may be driven by day-to-day variability in the NIRS breakpoints themselves or in the reference threshold used to validate them.

While the day-to-day variability of ventilatory and lactate thresholds has been extensively studied, the test-retest reliability of NIRS breakpoints remains largely unknown. Furthermore, although a wide range of NIRS breakpoint detection methods has been proposed, most have only been evaluated for criterion validity, not reliability.

Therefore, this exploratory analysis applies multiple breakpoint detection methods to a pre-existing data set of repeated incremental cycling tests and reports their test-retest reliability. The results should reveal differences between methods and inform future work on criterion validity and physiological interpretation of NIRS breakpoints.

## Sample

## Data Preprocessing

## Breakpoint methods applied

Methods for BP1:

1. Visual identification
2. double segmented on the full data (raw)
3. double segmented on the stage mean SmO~2~
4. double segmented on the slope per stage
5. double segmented on the power/SmO~2~ ratio
6. log-log method on stage mean SmO~2~
7. log-log method on the power/SmO~2~ ratio
8. second derivative of 4th degree polynomial (1st inflection)

Methods for BP2:

1. Visual identification
2. double segmented on the full data (raw)
3. double segmented on the stage mean SmO~2~
4. double segmented on the slope per stage
5. double segmented on the power/SmO~2~ ratio
6. Dmax method on stage mean SmO~2~
7. Dmax method on the power/SmO~2~ ratio
8. modDmax method on stage mean SmO~2~
9. modDmax method on power/SmO~2~ ratio
10. second derivative of 4th degree polynomial (2nd inflection)

## Load the data

The data was produced in the pre-processing pipeline, taking the raw recordings, applying smoothing, filtering (with `{mnirs}`) and calculating the different breakpoints.

```{r load-data}
#| fig-width: 10
#| fig-height: 7

dat <- read.csv("../../Data/AnalysisData/breakpoint_data.csv") |>
  mutate(bp2_type = NA_character_)

ratings <- read.csv("../../Data/ManualRatings/Rater_A_ratings.csv") |>
  mutate(
    method  = paste0("visual_", rater),
    warning = ifelse(!identifiable, "not_identifiable", NA_character_)
  ) |>
  select(session_id, channel, method, bp, bp_x, warning, power, bp2_type)

dat <- bind_rows(dat, ratings) |>
  mutate(
    id   = sub("_.*", "", session_id),
    test = sub(".*_", "", session_id)
  )

dat |>
    ggplot(aes(power, method, color = id)) +
    geom_point() +
    facet_grid(bp ~ channel) +
    labs(title = "Raw breakpoint power by method, channel, and BP", x = "Power (W)", y = NULL)

# now get subsets for dom, ndom and avg

dat_dom <- dat |> 
  dplyr::filter(channel == "smo2_dom")

dat_ndom <- dat |> 
  dplyr::filter(channel == "smo2_ndom")

dat_avg <- dat |> 
  dplyr::filter(channel == "smo2_avg")

dat_dom_wide <- dat_dom |>
  pivot_wider(
    id_cols    = c(id, method, bp),
    names_from = test,
    values_from = power,
    names_prefix = "T"
  ) |>
  mutate(
    mean_power = apply(pick(T01, T02, T03), 1, mean, na.rm = TRUE),
    sd_power   = apply(pick(T01, T02, T03), 1, sd,   na.rm = TRUE),
    cv_power   = sd_power / mean_power * 100,
    channel    = "DOM"
  )

dat_ndom_wide <- dat_ndom |>
  pivot_wider(
    id_cols    = c(id, method, bp),
    names_from = test,
    values_from = power,
    names_prefix = "T"
  ) |>
  mutate(
    mean_power = apply(pick(T01, T02, T03), 1, mean, na.rm = TRUE),
    sd_power   = apply(pick(T01, T02, T03), 1, sd,   na.rm = TRUE),
    cv_power   = sd_power / mean_power * 100,
    channel    = "NDOM"
  )

dat_avg_wide <- dat_avg |>
  pivot_wider(
    id_cols    = c(id, method, bp),
    names_from = test,
    values_from = power,
    names_prefix = "T"
  ) |>
  mutate(
    mean_power = apply(pick(T01, T02, T03), 1, mean, na.rm = TRUE),
    sd_power   = apply(pick(T01, T02, T03), 1, sd,   na.rm = TRUE),
    cv_power   = sd_power / mean_power * 100,
    channel    = "AVG"
  )

dat_all_wide <- bind_rows(dat_dom_wide, dat_ndom_wide, dat_avg_wide)
```

## Assumption testing

Before proceeding to get estimands for test-retest reliability, test some assumptions.

### Normality

Look at q-q plots and run a Shapiro-Wilk
```{r normality}
#| fig-width: 11
#| fig-height: 6

# for DOM
dat_dom |> 
  ggplot(aes(sample = power)) +
  stat_qq() +
  stat_qq_line() +
  facet_grid(bp ~ method) +
  labs(title = "QQ plots Breakpoints DOM")

dat_dom |> 
  group_by(method, bp) |> 
  rstatix::shapiro_test(power)

# for NDOM
dat_ndom |> 
  ggplot(aes(sample = power)) +
  stat_qq() +
  stat_qq_line() +
  facet_grid(bp ~ method) +
  labs(title = "QQ plots Breakpoints NDOM")

dat_ndom |> 
  group_by(method, bp) |> 
  rstatix::shapiro_test(power)
# for AVG
dat_avg |> 
  ggplot(aes(sample = power)) +
  stat_qq() +
  stat_qq_line() +
  facet_grid(bp ~ method) +
  labs(title = "QQ plots Breakpoints AVG")

dat_dom |> 
  group_by(method, bp) |> 
  rstatix::shapiro_test(power)
```

Both the inferential test and the q-q plots show that normality cannot be assumed for most methods. This is relevant for checking systematic trial effects on breakpoints. The violation of normality does not bias the ICC point estimate, but it affects the validity of its confidence interval — we therefore use bootstrapped CIs to quantify uncertainty.

### Homoscedasticity

Homoscedasticity of within-subject variability is assessed by plotting absolute pairwise differences against the mean threshold, for each method.

```{r homosced}
#| fig-width: 11
#| fig-height: 6

# for dom
# formal test using rank-order correlation
dat_dom_wide |>
  group_by(method, bp) |>
  filter(!is.na(T01) & !is.na(T02)) |>
  summarize(
    tryCatch(
      broom::tidy(
        cor.test(abs(T01 - T02), (T01 + T02) / 2,
        method = "spearman", exact = FALSE)
      ) |> select(estimate, p.value),
      error = function(e) tibble(estimate = NA_real_, p.value = NA_real_)
    ),
    n = n(),
    .groups = "drop"
  ) |>
  select(method, bp, rho = estimate, p.value, n)

  # Visual (primary evidence) — faceted across all 12 panels
dat_dom_wide |>
  filter(!is.na(T01), !is.na(T02)) |>
  mutate(
    abs_diff = abs(T01 - T02),
    mean_val = (T01 + T02) / 2,
    bp = factor(bp, levels = c(1, 2), labels = c("BP1", "BP2"))
  ) |>
  ggplot(aes(x = mean_val, y = abs_diff)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  facet_grid(bp ~ method) +
  labs(x = "Mean (T01+T02)/2", y = "|T01 - T02|",
       title = "Homoscedasticity check DOM")

# for ndom
# formal test using rank-order correlation
dat_ndom_wide |>
  group_by(method, bp) |>
  filter(!is.na(T01) & !is.na(T02)) |>
  summarize(
    tryCatch(
      broom::tidy(
        cor.test(abs(T01 - T02), (T01 + T02) / 2,
        method = "spearman", exact = FALSE)
      ) |> select(estimate, p.value),
      error = function(e) tibble(estimate = NA_real_, p.value = NA_real_)
    ),
    n = n(),
    .groups = "drop"
  ) |>
  select(method, bp, rho = estimate, p.value, n)

# Visual (primary evidence) — faceted across all 12 panels
dat_ndom_wide |>
  filter(!is.na(T01), !is.na(T02)) |>
  mutate(
    abs_diff = abs(T01 - T02),
    mean_val = (T01 + T02) / 2,
    bp = factor(bp, levels = c(1, 2), labels = c("BP1", "BP2"))
  ) |>
  ggplot(aes(x = mean_val, y = abs_diff)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  facet_grid(bp ~ method) +
  labs(x = "Mean (T01+T02)/2", y = "|T01 - T02|",
       title = "Homoscedasticity check NDOM")

# for avg
# formal test using rank-order correlation
dat_avg_wide |>
  group_by(method, bp) |>
  filter(!is.na(T01) & !is.na(T02)) |>
  summarize(
    tryCatch(
      broom::tidy(
        cor.test(abs(T01 - T02), (T01 + T02) / 2,
        method = "spearman", exact = FALSE)
      ) |> select(estimate, p.value),
      error = function(e) tibble(estimate = NA_real_, p.value = NA_real_)
    ),
    n = n(),
    .groups = "drop"
  ) |>
  select(method, bp, rho = estimate, p.value, n)

# Visual (primary evidence) — faceted across all 12 panels
dat_avg_wide |>
  filter(!is.na(T01), !is.na(T02)) |>
  mutate(
    abs_diff = abs(T01 - T02),
    mean_val = (T01 + T02) / 2,
    bp = factor(bp, levels = c(1, 2), labels = c("BP1", "BP2"))
  ) |>
  ggplot(aes(x = mean_val, y = abs_diff)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  facet_grid(bp ~ method) +
  labs(x = "Mean (T01+T02)/2", y = "|T01 - T02|",
       title = "Homoscedasticity check AVG")
```

No consistent pattern of heteroscedasticity emerges across the many comparisons. Rank-order correlations between the mean and the absolute difference tend to be low for DOM but higher for NDOM and AVG, possibly driven by outliers. Given this ambiguity, we report both SD and CV as complementary measures of variability for all methods.

### Systematic trial effect

With non-normal data, simply use Friedman test to look at order effects.

```{r order-effect}

# for DOM
dat_dom |>
  group_by(method, bp, id) |>
  filter(n() == 3, !any(is.na(power))) |>
  group_by(method, bp) |>
  nest() |>
  mutate(result = map(data, \(d) {
    tryCatch({
      res <- friedman.test(power ~ test | id, data = d)
      tibble(statistic = res$statistic, df = res$parameter, p.value = res$p.value)
    }, error = function(e) {
      tibble(statistic = NA_real_, df = NA_integer_, p.value = NA_real_)
    })
  })) |>
  select(-data) |>
  unnest(result)

# for NDOM
dat_ndom |>
  group_by(method, bp, id) |>
  filter(n() == 3, !any(is.na(power))) |>
  group_by(method, bp) |>
  nest() |>
  mutate(result = map(data, \(d) {
    tryCatch({
      res <- friedman.test(power ~ test | id, data = d)
      tibble(statistic = res$statistic, df = res$parameter, p.value = res$p.value)
    }, error = function(e) {
      tibble(statistic = NA_real_, df = NA_integer_, p.value = NA_real_)
    })
  })) |>
  select(-data) |>
  unnest(result)

# for AVG
dat_avg |>
  group_by(method, bp, id) |>
  filter(n() == 3, !any(is.na(power))) |>
  group_by(method, bp) |>
  nest() |>
  mutate(result = map(data, \(d) {
    tryCatch({
      res <- friedman.test(power ~ test | id, data = d)
      tibble(statistic = res$statistic, df = res$parameter, p.value = res$p.value)
    }, error = function(e) {
      tibble(statistic = NA_real_, df = NA_integer_, p.value = NA_real_)
    })
  })) |>
  select(-data) |>
  unnest(result)
```

The Friedman tests return non-significant effects for all methods. Thus, we conclude there are no relevant order (trial) effects on the breakpoints.

## Breakpoint differences between sides

First, compare the power output at BP1 and BP2 between DOM, NDOM and AVG. Because normalitiy may not be assumed, compare via pairwise bootstrap.

```{r side-diff-pwr}
#| fig-width: 9
#| fig-height: 6

dat_diff <- dat |>
  mutate(channel = gsub("smo2_", "", channel)) |>   # smo2_dom → dom etc.
  pivot_wider(
    id_cols     = c(id, test, session_id, method, bp),
    names_from  = channel,
    values_from = c(power, bp_x)                    # keep both metrics wide
  )

set.seed(42)

side_diff_all <- dat_diff |>
  drop_na(power_dom, power_ndom, power_avg) |>
  group_by(method, bp) |>
  nest() |>
  mutate(result = map(data, \(d) {
    if (nrow(d) < 3) return(tibble(
      comparison = character(), n = integer(), mean_diff = numeric(),
      ci_lower = numeric(), ci_upper = numeric(), p_raw = numeric()
    ))
    tests <- list(
      dom_vs_ndom = boot_t_test(d$power_dom,  d$power_ndom, paired = TRUE, R = 9999),
      dom_vs_avg  = boot_t_test(d$power_dom,  d$power_avg,  paired = TRUE, R = 9999),
      ndom_vs_avg = boot_t_test(d$power_ndom, d$power_avg,  paired = TRUE, R = 9999)
    )
    tibble(
      comparison = names(tests),
      n          = nrow(d),
      mean_diff  = c(
        mean(d$power_dom  - d$power_ndom, na.rm = TRUE),
        mean(d$power_dom  - d$power_avg,  na.rm = TRUE),
        mean(d$power_ndom - d$power_avg,  na.rm = TRUE)
      ),
      ci_lower = map_dbl(tests, \(t) t$conf.int[1]),
      ci_upper = map_dbl(tests, \(t) t$conf.int[2]),
      p_raw    = map_dbl(tests, "p.value")
    )
  })) |>
  select(-data) |>
  unnest(result) |>
  mutate(sig = p_raw < 0.05)

side_diff_all

side_diff_all |>
   filter(n >= 12) |>
  mutate(
    p_label = case_when(
      is.na(p_raw)  ~ "",
      p_raw < 0.001 ~ "p < .001",
      TRUE          ~ paste0("p = ", round(p_raw, 3))
    )
  ) |>
  ggplot(aes(mean_diff, method)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_linerange(aes(xmin = ci_lower, xmax = ci_upper, color = sig)) +
  geom_point(aes(color = sig)) +
  geom_text(aes(x = Inf, label = p_label), hjust = 1.05, size = 2.8) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "#e63946"),
                     labels = c("ns", "p < .05")) +
  facet_wrap(bp ~ comparison) +
  labs(x = "Mean difference (W)", y = NULL, color = NULL) +
    theme(legend.position = "bottom")

```

## Reliability quantification per method

### Failed breakpoint identification

A total of **36** tests were included (12 participants * 3 tests).

First, look at the number of breakpoints that could not be identified by method.

```{r bp-failed}
#| fig-width: 8
#| fig-height: 6

bp_failed <- bind_rows(
  dat_dom  |> group_by(method, bp) |> summarize(n_total = n(), failed = sum(is.na(power)), .groups = "drop") |> mutate(channel = "DOM"),
  dat_ndom |> group_by(method, bp) |> summarize(n_total = n(), failed = sum(is.na(power)), .groups = "drop") |> mutate(channel = "NDOM"),
  dat_avg  |> group_by(method, bp) |> summarize(n_total = n(), failed = sum(is.na(power)), .groups = "drop") |> mutate(channel = "AVG")
) |>
  mutate(
    pct_failed  = round(failed / n_total * 100),
    low_success = (100 - pct_failed) < 33
  )

exclude_combos <- bp_failed |>
  filter(low_success) |>
  select(method, channel, bp)

bp_failed |>
  mutate(
    bp    = factor(bp, labels = c("BP1", "BP2")),
    label = paste0(failed, "\n(", pct_failed, "%)")
  ) |>
  ggplot(aes(x = channel, y = method, fill = pct_failed)) +
  geom_tile(color = "white") +
  geom_text(aes(label = label), size = 2.8, lineheight = 0.9) +
  scale_fill_gradient(low = "#f0fdf4", high = "#ef4444",
                      name = "% failed", limits = c(0, 100)) +
  facet_wrap(~ bp) +
  labs(x = NULL, y = NULL, title = "Failed breakpoint estimates per method")


```

Then, look which type of visual BP2s were identified.

```{r bp2-visual-type}
#| fig-width: 6
#| fig-height: 4

bp2_visual <- dat |>
  filter(!is.na(bp2_type)) |>
  group_by(channel, bp2_type) |>
  summarize(n = n(), .groups = "drop")

bp2_visual |> 
  group_by(channel) |>
  mutate(pct = n / sum(n)) |>
  ggplot(aes(x = channel, y = pct, fill = bp2_type)) +
  geom_col() +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("flat" = "#60a5fa", "deoxy" = "#f87171")) +
  labs(x = NULL, y = NULL, fill = "BP2 type",
       title = "Visual BP2 type by channel")

```

### Reliability statistics

Calculate ICC(2,1) with 95% CI for each method. Select absolute agreement, two-way random effects, single measures.

```{r ICC}
#| fig-width: 10
#| fig-height: 9

icc_dom <- dat_dom_wide |>
  group_by(method, bp) |>
  nest() |>
  mutate(result = map(data, \(d) {
    tryCatch({
      res <- SimplyAgree::reli_stats(
        data = d, wide = TRUE, col.names = c("T01", "T02", "T03")
      )
      tibble(
        n         = res$n.id,
        icc       = res$icc$icc[res$icc$type == "ICC2"],
        icc_lower = res$icc$lower.ci[res$icc$type == "ICC2"],
        icc_upper = res$icc$upper.ci[res$icc$type == "ICC2"],
        sem       = res$SEM$estimate,
        cv        = res$cv$estimate,
        mdc95     = res$SEM$estimate * 1.96 * sqrt(2)
      )
    }, error = function(e) {
      tibble(n = NA_integer_, icc = NA_real_, icc_lower = NA_real_, icc_upper = NA_real_,
             sem = NA_real_, cv = NA_real_, mdc95 = NA_real_)
    })
  })) |>
  select(-data) |>
  unnest(result) |>
  mutate(channel = "DOM")

icc_ndom <- dat_ndom_wide |>
  group_by(method, bp) |>
  nest() |>
  mutate(result = map(data, \(d) {
    tryCatch({
      res <- SimplyAgree::reli_stats(
        data = d, wide = TRUE, col.names = c("T01", "T02", "T03")
      )
      tibble(
        n         = res$n.id,
        icc       = res$icc$icc[res$icc$type == "ICC2"],
        icc_lower = res$icc$lower.ci[res$icc$type == "ICC2"],
        icc_upper = res$icc$upper.ci[res$icc$type == "ICC2"],
        sem       = res$SEM$estimate,
        cv        = res$cv$estimate,
        mdc95     = res$SEM$estimate * 1.96 * sqrt(2)
      )
    }, error = function(e) {
      tibble(n = NA_integer_, icc = NA_real_, icc_lower = NA_real_, icc_upper = NA_real_,
             sem = NA_real_, cv = NA_real_, mdc95 = NA_real_)
    })
  })) |>
  select(-data) |>
  unnest(result) |>
  mutate(channel = "NDOM")

icc_avg <- dat_avg_wide |>
  group_by(method, bp) |>
  nest() |>
  mutate(result = map(data, \(d) {
    tryCatch({
      res <- SimplyAgree::reli_stats(
        data = d, wide = TRUE, col.names = c("T01", "T02", "T03")
      )
      tibble(
        n         = res$n.id,
        icc       = res$icc$icc[res$icc$type == "ICC2"],
        icc_lower = res$icc$lower.ci[res$icc$type == "ICC2"],
        icc_upper = res$icc$upper.ci[res$icc$type == "ICC2"],
        sem       = res$SEM$estimate,
        cv        = res$cv$estimate,
        mdc95     = res$SEM$estimate * 1.96 * sqrt(2)
      )
    }, error = function(e) {
      tibble(n = NA_integer_, icc = NA_real_, icc_lower = NA_real_, icc_upper = NA_real_,
             sem = NA_real_, cv = NA_real_, mdc95 = NA_real_)
    })
  })) |>
  select(-data) |>
  unnest(result) |>
  mutate(channel = "AVG")

icc_all <- bind_rows(icc_dom, icc_ndom, icc_avg)

# plot the results ordered by icc
icc_all <- icc_all |>
  group_by(bp) |>
  arrange(desc(icc)) |>
  ungroup() |>
  mutate(combination = interaction(channel, method, sep = "_"))

dir.create("../../Data/AnalysisData", showWarnings = FALSE, recursive = TRUE)
saveRDS(icc_all, "../../Data/AnalysisData/icc_all.rds")

icc_all |>
  # filter(n >= 12) |>
  anti_join(exclude_combos, by = c("method", "channel", "bp")) |>
  ggplot(aes(x = icc, y = tidytext::reorder_within(combination, icc, bp), color = channel)) +
  geom_point() +
  geom_pointrange(aes(xmin = icc_lower, xmax = icc_upper)) +
  tidytext::scale_y_reordered() +
  facet_wrap(~ bp, scales = "free_y") +
  labs(x = "ICC(2,1)", y = NULL) +
  theme(legend.position = "bottom")

icc_all |>
  # filter(n >= 12) |>
  anti_join(exclude_combos, by = c("method", "channel", "bp")) |>
  ggplot(aes(x = sem, y = tidytext::reorder_within(combination, -sem, bp), fill = channel)) +
  geom_bar(stat = "identity") +
  tidytext::scale_y_reordered() +
  facet_wrap(~ bp, scales = "free_y") +
  labs(x = "SEM [W]", y = NULL) +
  theme(legend.position = "bottom")

icc_all |>
  # filter(n >= 12) |>
  anti_join(exclude_combos, by = c("method", "channel", "bp")) |>
  ggplot(aes(x = cv, y = tidytext::reorder_within(combination, -cv, bp), fill = channel)) +
  geom_bar(stat = "identity") +
  tidytext::scale_y_reordered() +
  scale_x_continuous(labels = percent) +
  facet_wrap(~ bp, scales = "free_y") +
  labs(x = "CV", y = NULL) +
  theme(legend.position = "bottom")
```

Calculate SEM (√MS_error from the two-way repeated measures ANOVA).

Calculate MDC₉₅ (SEM × 1.96 × √2).

## Comparing reliability between methods

Run the paired bootstrap for each method pair: resample the 12 participant-level pairs with replacement, compute the mean difference in within-subject SD and CV each iteration (10,000 resamples), derive the 95% CI. If the CI excludes zero, conclude the methods differ in reliability.

```{r bootstrap-reli}
#| fig-width: 12
#| fig-height: 16

set.seed(42)

boot_reli <- dat_all_wide |>
  group_by(channel, bp) |>
  nest() |>
  mutate(result = map(data, \(d) {
    pairs <- combn(unique(d$method), 2, simplify = FALSE)
    map_dfr(pairs, \(pair) {
      d_wide <- d |>
        filter(method %in% pair) |>
        pivot_wider(id_cols = id, names_from = method,
                    values_from = c(sd_power, cv_power))

      sd1 <- d_wide[[paste0("sd_power_", pair[1])]]
      sd2 <- d_wide[[paste0("sd_power_", pair[2])]]
      cv1 <- d_wide[[paste0("cv_power_", pair[1])]]
      cv2 <- d_wide[[paste0("cv_power_", pair[2])]]

      run_boot <- function(x, y) {
        ok <- !is.na(x) & !is.na(y)
        if (sum(ok) < 3) return(tibble(n = sum(ok), mean_diff = NA_real_,
                                       ci_lower = NA_real_, ci_upper = NA_real_,
                                       p_raw = NA_real_))
        res <- boot_t_test(x[ok], y[ok], paired = TRUE, R = 9999)
        tibble(n = sum(ok), mean_diff = mean(x[ok] - y[ok]),
               ci_lower = res$conf.int[1], ci_upper = res$conf.int[2],
               p_raw = res$p.value)
      }

      bind_rows(
        run_boot(sd1, sd2) |> mutate(metric = "SD"),
        run_boot(cv1, cv2) |> mutate(metric = "CV")
      ) |> mutate(method1 = pair[1], method2 = pair[2])
    })
  })) |>
  select(-data) |>
  unnest(result) |>
  mutate(
    sig  = p_raw < 0.05,
    pair = paste0(method1, " vs ", method2)
  )

# ── Figure: SD ────────────────────────────────────────────────────────────────
boot_reli |>
  filter(metric == "SD", !is.na(mean_diff)) |>
  anti_join(rename(exclude_combos, method1 = method), by = c("method1", "channel", "bp")) |>
  anti_join(rename(exclude_combos, method2 = method), by = c("method2", "channel", "bp")) |>
  mutate(
    p_label = case_when(
      p_raw < 0.001 ~ "p < .001",
      TRUE          ~ paste0("p = ", round(p_raw, 3))
    ),
    bp = factor(bp, levels = c(1, 2), labels = c("BP1", "BP2"))
  ) |>
  ggplot(aes(x = mean_diff, y = reorder_within(pair, mean_diff, list(channel, bp)))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_linerange(aes(xmin = ci_lower, xmax = ci_upper, color = sig)) +
  geom_point(aes(color = sig)) +
  geom_text(aes(x = Inf, label = p_label), hjust = 1.05, size = 2.5) +
  scale_y_reordered() +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "#e63946"),
                     labels = c("ns", "p < .05")) +
  facet_grid(bp ~ channel, scales = "free_y", space = "free") +
  labs(x = "Mean difference in SD (W)", y = NULL, color = NULL,
       title = "Pairwise reliability comparison — SD") +
    theme(legend.position = "bottom")

# ── Figure: CV ────────────────────────────────────────────────────────────────
boot_reli |>
  filter(metric == "CV", !is.na(mean_diff)) |>
  anti_join(rename(exclude_combos, method1 = method), by = c("method1", "channel", "bp")) |>
  anti_join(rename(exclude_combos, method2 = method), by = c("method2", "channel", "bp")) |>
  mutate(
    p_label = case_when(
      p_raw < 0.001 ~ "p < .001",
      TRUE          ~ paste0("p = ", round(p_raw, 3))
    ),
    bp = factor(bp, levels = c(1, 2), labels = c("BP1", "BP2"))
  ) |>
  ggplot(aes(x = mean_diff, y = reorder_within(pair, mean_diff, list(channel, bp)))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  geom_linerange(aes(xmin = ci_lower, xmax = ci_upper, color = sig)) +
  geom_point(aes(color = sig)) +
  geom_text(aes(x = Inf, label = p_label), hjust = 1.05, size = 2.5) +
  scale_y_reordered() +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "#e63946"),
                     labels = c("ns", "p < .05")) +
  facet_grid(bp ~ channel, scales = "free_y", space = "free") +
  labs(x = "Mean difference in CV (%)", y = NULL, color = NULL,
       title = "Pairwise reliability comparison — CV") +
    theme(legend.position = "bottom")

```


## Summary

### Detection failure rates

```{r summary-detection-table}
bp_failed |>
  mutate(
    bp    = factor(bp, levels = c(1, 2), labels = c("BP1", "BP2")),
    label = paste0(failed, "/", n_total, " (", pct_failed, "%)")
  ) |>
  group_by(method, channel) |>
  mutate(sort_key = mean(pct_failed)) |>
  ungroup() |>
  pivot_wider(
    id_cols    = c(method, channel, sort_key),
    names_from  = bp,
    values_from = label,
    values_fill = "0/0 (0%)"
  ) |>
  arrange(sort_key) |>
  select(-sort_key) |>
  knitr::kable(
    col.names = c("Method", "Channel", "BP1", "BP2"),
    caption   = "Table S1. Failed breakpoint detections per method and channel."
  )
```

### Reliability statistics

Main manuscript table: ICC(2,1) with 95% CI, SEM, CV, and MDC₉₅ per method and channel.
Groups with fewer than 7 observations are excluded.

```{r summary-reli-table}
icc_all |>
  # filter(n >= 8) |>
  anti_join(exclude_combos, by = c("method", "channel", "bp")) |>
  mutate(
    BP     = factor(bp, levels = c(1, 2), labels = c("BP1", "BP2")),
    icc_ci = sprintf("%.2f [%.2f, %.2f]", icc, icc_lower, icc_upper),
    sem_r  = round(sem, 1),
    cv_pct = round(cv * 100, 1),
    mdc_r  = round(mdc95, 1)
  ) |>
  arrange(BP, sem) |>
  select(
    Method   = method,
    Channel  = channel,
    BP,
    n,
    `ICC(2,1) [95% CI]` = icc_ci,
    `SEM (W)`           = sem_r,
    `CV (%)`            = cv_pct,
    `MDC₉₅ (W)`        = mdc_r
  ) |>
  knitr::kable(caption = "Table 1. Test-retest reliability per method and channel (n >= 12).")
```

### Channel differences (side comparisons)

Significant pairwise bootstrap comparisons between DOM, NDOM, and AVG.

```{r summary-side-diffs}
exclude_method_bp  <- exclude_combos |> distinct(method, bp)
side_diff_filtered <- side_diff_all |>
  filter(!is.na(p_raw)) |>
  anti_join(exclude_method_bp, by = c("method", "bp"))
n_side_total <- nrow(side_diff_filtered)
n_side_sig   <- sum(side_diff_filtered$sig)

side_diff_filtered |>
  filter(sig) |>
  mutate(
    bp      = factor(bp, levels = c(1, 2), labels = c("BP1", "BP2")),
    result  = sprintf("Δ = %.1f W [%.1f, %.1f], p = %.3f", mean_diff, ci_lower, ci_upper, p_raw)
  ) |>
  arrange(desc(abs(mean_diff))) |>
  select(
    Method     = method,
    BP         = bp,
    Comparison = comparison,
    n,
    Result     = result
  ) |>
  knitr::kable(
    caption = sprintf(
      "Table S2. Significant channel differences in breakpoint power output (%d of %d comparisons significant, n >= 12).",
      n_side_sig, n_side_total
    )
  )
```

### Method reliability comparisons

Significant pairwise bootstrap comparisons of within-subject SD and CV between methods.

```{r summary-method-comparisons}
boot_reli_filtered <- boot_reli |>
  filter(!is.na(p_raw)) |>
  anti_join(rename(exclude_combos, method1 = method), by = c("method1", "channel", "bp")) |>
  anti_join(rename(exclude_combos, method2 = method), by = c("method2", "channel", "bp"))

fmt_reli <- function(data) {
  data |>
    mutate(
      bp     = factor(bp, levels = c(1, 2), labels = c("BP1", "BP2")),
      result = sprintf("Δ = %.2f [%.2f, %.2f], p = %.3f", mean_diff, ci_lower, ci_upper, p_raw)
    ) |>
    arrange(desc(abs(mean_diff))) |>
    select(Channel = channel, BP = bp, Comparison = pair, n, Result = result)
}

# SD
sd_filtered <- boot_reli_filtered |> filter(metric == "SD")
fmt_reli(filter(sd_filtered, sig)) |>
  knitr::kable(
    caption = sprintf(
      "Table S3a. Significant pairwise differences in within-subject SD between methods (%d of %d comparisons significant, n >= 12).",
      sum(sd_filtered$sig), nrow(sd_filtered)
    )
  )

# CV
cv_filtered <- boot_reli_filtered |> filter(metric == "CV")
fmt_reli(filter(cv_filtered, sig)) |>
  knitr::kable(
    caption = sprintf(
      "Table S3b. Significant pairwise differences in CV between methods (%d of %d comparisons significant, n >= 12).",
      sum(cv_filtered$sig), nrow(cv_filtered)
    )
  )
```

## Discussion

Frame conclusions carefully given n=12: distinguish between "method A showed better reliability" (descriptive) and "methods differed significantly in reliability" (inferential). With this sample size, absence of a significant difference is not equivalence.

## Future analyses

- **Bland-Altman plots (repeated-measures extension, Bland & Altman 2007).** With 3 trials there are 3 pairwise comparisons per participant (T1–T2, T1–T3, T2–T3); the repeated-measures version handles the non-independence of these pairs correctly. Would report bias and LoA with their own 95% CIs.

## References